Chemical tagging can work: \\ Identification of stellar phase-space structures \\ purely by chemical-abundance similarity

### 2016-01-19 papers/clusters.tex
text tweaks, no footnote

% to-do
% - submit to arXiv
% - forward to Freeman, Bovy, others?
% - submit to journal

@@

In what follows, we are going to use \apogee\ \acronym{DR12} data \citep{dr12}, in
which we can re-derive 15 element abundances (C, N, O, Na, Mg, Al, Si,
S, K, Ca, Ti, V, Mn, Fe, Ni), using \thecannon\ \citep{thecannon}.
The \apogee\ data set covers a huge radial extent and---because the
data are taken in the infrared---explores the thin disk.
However, it has the disadvantage that it's spatial coverage is incomplete
(i.e., limited pointings produce a \foreign{`swiss cheese'}-like structure),
which makes it hard to see, within the data set, linear or
extended stellar structures.
In many ways, \galah\ will deliver improvements, both because it will
have more abundances (possibly 29) and contiguous sky coverage.

@@

Another issue with k-means is that it effectively uses metric
distances in the $D$-space; it presumes Euclidean isotropy.
We choose here to work in the hydrogen-normalized abundance space, the
space of [C/H], [N/H], [O/H], [Na/H], [Mg/H], [Al/H], [Si/H], [S/H],
[K/H], [Ca/H], [Ti/H], [V/H], [Mn/H], [Fe/H], [Ni/H].
But in addition to this, we re-scale these by approximate typical measurement
precisions obtained by \thecannon\ before running k-means.\footnote{%
  The scalings were
  [C/H]/[0.041],
  [N/H]/[0.044],
  [O/H]/[0.037],
  [Na/H]/[0.111],
  [Mg/H]/[0.032],
  [Al/H]/[0.055],
  [Si/H]/[0.041],
  [S/H]/[0.054],
  [K/H]/[0.069],
  [Ca/H]/[0.043],
  [Ti/H]/[0.072],
  [V/H]/[0.146],
  [Mn/H]/[0.041],
  [Fe/H]/[0.019], and
  [Ni/H]/[0.034].}
This makes the space close to isotropic in measurement uncertainty or
observational precision.
Importantly we use \emph{only} abundance-space information, and no
positional or velocity information (nor $\teff$ nor $\logg$ nor any
targeting or observational meta-data) as input to the clustering
algorithm.

>>>>

@@

In what follows, we are going to use \apogee\ \acronym{DR12} data \citep{dr12}, in
which we can re-derive 15 element abundances (C, N, O, Na, Mg, Al, Si,
S, K, Ca, Ti, V, Mn, Fe, Ni), using \thecannon\ \citep{thecannon}.
The \apogee\ data set covers a huge radial extent and---because the
data are taken in the infrared---explores the thin disk.
However, it has the disadvantage that it's spatial coverage is incomplete
(that is, limited pointings produce a fractured sky map),
which makes it hard to see, within the data set, linear or
extended stellar structures.
In many ways, \galah\ will deliver improvements, both because it will
have more abundances (possibly 29) and contiguous sky coverage.

@@

Another issue with k-means is that it effectively uses metric
distances in the $D$-space; it presumes Euclidean isotropy.
We choose here to work in the hydrogen-normalized abundance space, the
space of [C/H], [N/H], [O/H], [Na/H], [Mg/H], [Al/H], [Si/H], [S/H],
[K/H], [Ca/H], [Ti/H], [V/H], [Mn/H], [Fe/H], [Ni/H].
But in addition to this, we re-scale these by approximate typical measurement
precisions obtained by \thecannon\ before running k-means.
These scalings were
  [C/H]/[0.041],
  [N/H]/[0.044],
  [O/H]/[0.037],
  [Na/H]/[0.111],
  [Mg/H]/[0.032],
  [Al/H]/[0.055],
  [Si/H]/[0.041],
  [S/H]/[0.054],
  [K/H]/[0.069],
  [Ca/H]/[0.043],
  [Ti/H]/[0.072],
  [V/H]/[0.146],
  [Mn/H]/[0.041],
  [Fe/H]/[0.019], and
  [Ni/H]/[0.034].
This scaling makes the space close to isotropic in measurement uncertainty or
observational precision.
Importantly we use \emph{only} abundance-space information, and no
positional or velocity information (nor $\teff$ nor $\logg$ nor any
targeting or observational meta-data) as input to the clustering
algorithm.
